A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-17 papers/static-checker/ecoop2024/ecoop2024.tex
appendix
change the figures (kolibri) and (aution)

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.1\textwidth]{auction-life-cycle}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

@@

In a blockchain ecosystem, tokens represent digital assets that can be
transferred between accounts. These tokens are securely stored and
verified within the blockchain network. There are various types of
tokens, including native tokens (cryptocurrencies) like Bitcoin,
Ether, Tezos, and others. Additionally, tokens can be created as
overlays through smart contracts on an existing blockchain. Two common
categories are: (1) Fungible Tokens (FT), which are interchangeable
and (2) Non-Fungible Tokens (NFTs), which represent unique digital
assets, and each unit is distinct. Stablecoins, falling under fungible
tokens, are designed to maintain a stable value by tracking the value
of a fiat currency like USD or EUR.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) 
   && get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) 
   && get_big_map (from, ledger) = Some sbvar 
   && sbvar = Pair sbvar1 sbvar2 
   && not (sbvar1 >= value) : 'NotEnoughBalance'
4. not (paused = true) 
   && get_big_map (from, ledger) = Some sbvar 
   && sbvar = Pair sbvar1 sbvar2 
   && get_map (Sender, sbvar2) = Some sbvar3 
   && sbvar1 >= value 
   && not (sbvar3 >= value): 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
 storage :=  Pair (ledger: big_map address 
  (pair nat (map address nat))) 
  (Pair (admin: address) 
          (Pair (paused: bool) (total_supply: nat)))
 entrypoint %transfer
   parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
   pre-condition := not (paused  = true)
     && get_big_map (from, ledger) = 
       Some (v: pair nat (map address nat))
     && v = Pair (n: nat) (m: map address nat)
     && get_map (Sender, m) = Some (k: nat) 
     && (n >= value) && (k >= value)
   post-condition := get_big_map (from, post_ledger) = 
     Some (post_v:  pair nat (map address nat))
     && post_v = Pair (post_n: nat) (post_m: map address nat) 
     && get_map (Sender, post_m) = Some (post_k: nat) 
     && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

As an example, we demonstrate how to verify the \lstinline/getBalance/
entrypoint,  shown in Listing~\ref{lst:specification-get-balance}. The
other entrypoints \lstinline/getAllowance/ and
\lstinline/getTotalSupply/ can be specified similarly.  There are two
cases to consider: (1) if the \lstinline/owner/ is in the
\lstinline/ledger/ big map and (2) if it is not. For the former case,
lines 5-8 specify that \lstinline/owner/ is in the \lstinline/ledger/
big map, as the \lstinline/get_big_map/ function returns a value
\lstinline/Some value/ of the \TOPTION\ type, and this
\lstinline/value/ is a pair of a natural number \lstinline/k/ and a
map \lstinline/m/. The postcondition guarantees that the storage
remains unchanged, as specified in lines 10-13, and the call returns a
\lstinline/TRANSFER_TOKENS/ operation with the \lstinline/k/ number
and \lstinline/Amount/ to be passed to address \lstinline/c/. The
precondition in lines 16-17 specifies the latter case when the
\lstinline/owner/ does not exist in the big map \lstinline/ledger/; in
this case, the number \lstinline/0/ should be sent, and again, the
storage should remain unchanged.

@@

The specification of  the contract is showned in Listing \ref{lst:kolibri-contract-specification}. Here, we shortened the names of the variables for the storage.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{kolibri}
    \caption{Kolibri oracle entrypoint relations}
    \label{fig:kolibri-oracle-emtrypoint-relations}
\end{figure}

@@

    post-condition := state = 1

@@

    post-condition := state = 0 

@@

    post-condition := (governor = gover) 

@@

    post-condition := (maxDelay = delay) 

@@

&& get_contract (Nat, sbvar) = None: Unit

@@

\bibliography{ecoop2024}
\include{appendix}
\end{document}

>>>>

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.9\textwidth]{auction}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

@@

In a blockchain ecosystem, tokens represent digital assets that can be
transferred between accounts. These tokens are securely stored and
verified within the blockchain network. There are various types of
tokens, including native tokens (cryptocurrencies) like Bitcoin,
Ether, Tezos, and others. Additionally, tokens can be created as
overlays through smart contracts on an existing blockchain. Two common
categories are: (1) Fungible Tokens (FTs), which are interchangeable
and (2) Non-Fungible Tokens (NFTs), which represent unique digital
assets, and each unit is distinct. Stablecoins, falling under fungible
tokens, are designed to maintain a stable value by tracking the value
of a fiat currency like USD or EUR.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) 
   && get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) 
   && get_big_map (from, ledger) = Some sbvar 
   && sbvar = Pair sbvar1 sbvar2 
   && not (sbvar1 >= value) : 'NotEnoughBalance'
4. not (paused = true) 
   && get_big_map (from, ledger) = Some sbvar 
   && sbvar = Pair sbvar1 sbvar2 
   && get_map (Sender, sbvar2) = None: 'NotEnoughAllowance'
5. not (paused = true) 
   && get_big_map (from, ledger) = Some sbvar 
   && sbvar = Pair sbvar1 sbvar2 
   && get_map (Sender, sbvar2) = Some sbvar3 
   && sbvar1 >= value 
   && not (sbvar3 >= value): 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
 storage :=  Pair (ledger: big_map address 
  (pair nat (map address nat))) 
  (Pair (admin: address) 
          (Pair (paused: bool) (total_supply: nat)))
 entrypoint %transfer
   parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
   pre-condition := not (paused  = true)
     && get_big_map (from, ledger) = 
       Some (v: pair nat (map address nat))
     && v = Pair (n: nat) (m: map address nat)
     && get_map (Sender, m) = Some (k: nat) 
     && (n >= value) && (k >= value)
   post-condition := get_big_map (from, post_ledger) = 
     Some (post_v:  pair nat (map address nat))
     && post_v = Pair (post_n: nat) (post_m: map address nat) 
     && get_map (Sender, post_m) = Some (post_k: nat) 
     && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

As an example, we demonstrate how to verify the \lstinline/getBalance/
entrypoint,  shown in Listing~\ref{lst:specification-get-balance}. The
other entrypoints \lstinline/getAllowance/ and
\lstinline/getTotalSupply/ can be specified similarly.  There are two
cases to consider: (1) if the \lstinline/owner/ is in the
\lstinline/ledger/ big map and (2) if it is not. For the former case,
lines 3-5 specify that \lstinline/owner/ is in the \lstinline/ledger/
big map, as the \lstinline/get_big_map/ function returns a value
\lstinline/Some value/ of the \TOPTION\ type, and this
\lstinline/value/ is a pair of a natural number \lstinline/k/ and a
map \lstinline/m/. The postcondition guarantees that the storage
remains unchanged, as specified in lines 6-11, and the call returns a
\lstinline/TRANSFER_TOKENS/ operation with the \lstinline/k/ number
and \lstinline/Amount/ to be passed to address \lstinline/c/. The
precondition in lines 13-20 specifies the latter case when the
\lstinline/owner/ does not exist in the big map \lstinline/ledger/; in
this case, the number \lstinline/0/ should be sent, and again, the
storage should remain unchanged.

@@

The specification of  the contract is showned in Listing \ref{lst:kolibri-contract-specification}. Here, we shortened the names of the variables for the storage.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{kolibri(1)}
    \caption{Kolibri oracle entrypoint relations}
    \label{fig:kolibri-oracle-emtrypoint-relations}
\end{figure}

@@

    post-condition := post_state = 1

@@

    post-condition := post_state = 0 

@@

    post-condition := (post_governor = gover) 

@@

    post-condition := (post_maxDelay = delay) 

@@

&& get_contract (nat, sbvar) = None: Unit

@@

\bibliography{ecoop2024}
\appendix
\include{appendix}
\end{document}
